---
title: "Bayesian Final Exam"
author: "Bernard Gonzales"
date: "2024-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(ggplot2)
library(brms)
visagetome_data <- read.csv('DS6040F24_finaldat.csv')
```


We'll examine how all of the time and demographic variables affect the amount of time users spent on VisageTome using a hierarchical Bayesian model, and we will see how much individual variability plays a role in determining the amount of time on VisageTome.

```{r}
formula <- MinPerBlock ~ 1 + time_of_day + day_of_week + week + 
           gender + age + nationality + 
           (1 | user_id)

fit <- brm(
  formula = formula,
  data = visagetome_data,
  family = gaussian(),
  prior = c(
    prior(normal(0, 10), class = "Intercept"),
    prior(normal(0, 10), class = "b"),
    prior(cauchy(0, 2), class = "sd"),
    prior(cauchy(0, 2), class = "sigma")
  ),
  chains = 4, iter = 2000, cores = 4
)

summary(fit)
```

```{r, eval=FALSE}
> summary(fit)
 Family: gaussian 
  Links: mu = identity; sigma = identity 
Formula: MinPerBlock ~ 1 + time_of_day + day_of_week + week + gender + age + nationality + (1 | user_id) 
   Data: visagetome_data (Number of observations: 4200) 
  Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
         total post-warmup draws = 4000

Multilevel Hyperparameters:
~user_id (Number of levels: 100) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)    51.40      7.42    39.75    68.98 1.01      350      486

Regression Coefficients:
                        Estimate Est.Error l-95% CI u-95% CI  Rhat Bulk_ESS Tail_ESS
Intercept                   1.78     17.65   -35.57    33.93  1.00      587     1062
time_of_dayMidday         -13.43      3.50   -20.30    -6.67  1.00     5848     3006
time_of_dayMorning        -12.63      3.45   -19.44    -5.92  1.00     5157     3168
day_of_weekMonday         -20.12      4.63   -29.22   -11.21  1.00     4404     3060
day_of_weekSaturday        18.98      4.62    10.02    28.04  1.00     3496     3001
day_of_weekSunday         -33.95      4.63   -42.79   -24.82  1.00     4074     3193
day_of_weekThursday       -39.07      4.68   -48.39   -29.82  1.00     3811     3215
day_of_weekTuesday        -25.58      4.55   -34.43   -16.71  1.00     3892     2796
day_of_weekWednesday      -39.68      4.72   -48.80   -30.26  1.00     3697     2919
week                       -1.35      2.99    -7.16     4.53  1.00     6570     3105
genderMale                 -1.89      7.28   -15.80    12.50  1.00     1023     1849
age                         2.10      0.20     1.71     2.49  1.00      807     1311
nationalityCarthaginian     7.73      7.80    -7.71    23.11  1.00     1196     2227
nationalityFirstNations     5.89      7.97   -10.21    20.89  1.00     1889     2505
nationalityHolyRoman       -4.28      8.72   -21.07    13.15  1.00     2367     2768
nationalityRoman           -0.30      9.01   -17.50    17.65  1.00     2438     2690

Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma   101.35      1.11    99.27   103.55 1.00     7471     2921

Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).
```

The multilevel hyperparameter `sd(Intercept)` measures the variability of baseline usage across users. The value of 51.40 is pretty high, indicating that there is significant variation between individual users in their overall app usage.

The regression coefficients measure the effect of temporal and demographic variables on time spent on VisageTome. The `Intercept` coefficient represents the baseline app usage for a reference user who is female, age 18, and Atlantian, during the evening time block on Friday during the first week. The estimate of 1.78 is really small, compared to the 95% credible interval of [-35.57, 33.93], which means that this baseline is not well constrained.

Let's look at the effect of the temporal variables on VisageTome usage. Users spend 13 minutes less on VisageTome in the morning and midday than the evening. For day of the week, Saturdays had the most usage time, then Fridays, which averaged 19 minutes less than Saturdays. The rest of the days (Sunday through Thursday) had significantly less time spent on VisageTome. Weeks 1 and 2 had very little change, meaning that the week variable was statistically insignificant.

Next, let's see if demographic variables affected which users spent more or less time on VisageTome. Age is the only significant demographic predictor, and it indicates that older users spend more time on VisageTome than younger users. Gender and nationality had little to no effect on app usage.

The residual variance (measured by `sigma`) is the unexplained variability after accounting for the predictors. The estimate for `sigma` is 101.35, which means that there is substantial variability in app usage not captured by the model.

Let's investigate the possibility of superusers, who are users who spend too much time on VisageTome. We'll start by graphing the average (mean) of `MinPerBlock` per user, sorted by ascending order.

```{r}
user <- visagetome_data %>%
  group_by(user_id) %>%
  summarize(avg_screen_time = mean(MinPerBlock)) %>%
  arrange(avg_screen_time)
user$user_id = 1:100
ggplot(user, aes(x = user_id, y = avg_screen_time)) +
  geom_point(color = ifelse(user$avg_screen_time > 240, 'red', 'black')) +
  xlab('Users sorted from Least to Most Time on VisageTome')
```

As we can see, there are a lot of users who somehow use VisageTome for more than four hours within the given 4-hour blocks of time. In the graph, I colored the dots indicating superusers in red, while the rest of the users have black dots. I decided on the cutoff point being 240 minutes, or four hours, since the user would have to be using VisageTome on several devices at the same time in order to be able to have more than 240 minutes of time on VisageTome within a four hour period. I will add a `superuser` variable to the data, and run the model again to see if superuser status affects the usefulness of the model.

```{r}
user <- visagetome_data %>%
  group_by(user_id) %>%
  summarize(avg_screen_time = mean(MinPerBlock))
user$superuser <- ifelse(user$avg_screen_time > 240, 1, 0)
visagetome_data$superuser <- user$superuser

new_formula <- MinPerBlock ~ 1 + superuser + time_of_day + day_of_week + week + gender + age + nationality + 
               superuser:(gender + age + nationality) + (1 | user_id)

fit <- brm(
  formula = new_formula,
  data = visagetome_data,
  family = gaussian(),
  prior = c(
    prior(normal(0, 10), class = "Intercept"),
    prior(normal(0, 10), class = "b"),
    prior(cauchy(0, 2), class = "sd"),
    prior(cauchy(0, 2), class = "sigma")
  ),
  chains = 4, iter = 2000, cores = 4
)

summary(fit)
```

```{r, eval=FALSE}
> summary(fit)
 Family: gaussian 
  Links: mu = identity; sigma = identity 
Formula: MinPerBlock ~ 1 + superuser + time_of_day + day_of_week + week + gender + age + nationality + superuser:(gender + age + nationality) + (1 | user_id) 
   Data: visagetome_data (Number of observations: 4200) 
  Draws: 4 chains, each with iter = 2000; warmup = 1000; thin = 1;
         total post-warmup draws = 4000

Multilevel Hyperparameters:
~user_id (Number of levels: 100) 
              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sd(Intercept)    51.82      7.28    39.98    68.09 1.02      289      396

Regression Coefficients:
                                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
Intercept                             2.35     17.32   -33.07    34.20 1.01      607     1062
superuser                            -8.21      7.68   -23.65     6.91 1.00     4213     2873
time_of_dayMidday                   -13.50      3.46   -20.35    -6.54 1.00     4862     3052
time_of_dayMorning                  -12.62      3.55   -19.65    -5.71 1.00     6019     3056
day_of_weekMonday                   -20.20      4.61   -29.06   -11.23 1.00     3722     2968
day_of_weekSaturday                  18.80      4.59    10.04    27.52 1.00     4440     3112
day_of_weekSunday                   -34.00      4.66   -42.95   -24.62 1.00     3716     2989
day_of_weekThursday                 -39.14      4.62   -48.55   -29.98 1.00     3950     3162
day_of_weekTuesday                  -25.71      4.68   -35.13   -16.73 1.00     4616     3216
day_of_weekWednesday                -39.62      4.66   -48.65   -30.70 1.00     4008     2966
week                                 -1.54      2.95    -7.41     4.07 1.00     6163     2442
genderMale                           -1.41      7.41   -15.85    13.08 1.00     1325     1925
age                                   2.09      0.20     1.69     2.46 1.00      751     1317
nationalityCarthaginian               7.16      7.80    -8.61    21.92 1.00     1586     2201
nationalityFirstNations               5.72      8.01   -10.09    20.94 1.00     1932     2492
nationalityHolyRoman                 -4.34      8.65   -21.24    12.52 1.00     2384     2921
nationalityRoman                     -0.47      8.87   -17.73    16.88 1.00     2729     2841
superuser:genderMale                  0.06      6.47   -12.91    12.97 1.00     5937     2829
superuser:age                         0.08      0.12    -0.15     0.30 1.00     4180     2676
superuser:nationalityCarthaginian     4.57      7.04    -9.10    18.26 1.00     5551     2946
superuser:nationalityFirstNations     4.01      7.27   -10.61    18.29 1.00     5541     3163
superuser:nationalityHolyRoman       -1.90      7.93   -17.36    14.01 1.00     6745     3052
superuser:nationalityRoman           -1.43      8.65   -17.90    15.59 1.00     6403     3059


Further Distributional Parameters:
      Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
sigma   101.34      1.14    99.11   103.57 1.00     6054     2946

Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
and Tail_ESS are effective sample size measures, and Rhat is the potential
scale reduction factor on split chains (at convergence, Rhat = 1).
```

Most of the variables are the same as the previous model. What we are interested in is whether or not any generalizations can be made about superusers.













